\LARGE \bf Informed sampling by Markov Chain Monte Carlo methods

### 2017-09-12 paper.tex
updated pseudo code
update
updates on sec V - alg

\usepackage{xspace}
\usepackage{amsmath,amssymb,amsfonts,amsfonts}
\usepackage{algorithm}
\usepackage[noend]{algorithmic}
\usepackage{url}
\usepackage{bm}
\usepackage{graphicx}
\usepackage[font=footnotesize]{caption} % incompatible with aaai.sty?
\usepackage[font=footnotesize]{subcaption}
\usepackage{color}
\usepackage{cite}
\usepackage{hyperref}
%\hypersetup{bookmarksopen,
%bookmarksnumbered,
%pdfpagemode=UseOutlines,
%colorlinks=true,
%linkcolor=blue,
%anchorcolor=blue,
%citecolor=blue,
%filecolor=blue,
%menucolor=blue,
%urlcolor=blue
%}

@@

\section{MCMC-based Informed Sampling}
In this section we describe our approach to efficiently produce samples in the informed set~\Cinf given a specific cost~$c_{\text{best}}$.
Our algorithm (see Alg.~\ref{alg:mcmc_informed_sampling}) starts by randomly sampling a state~$x_\text{rand} \in \calX$ (line~1).
This sample is used to produces a 
sample~$x_0 \in \partial\Cinf$ 
%sample~$x_\text{boundary} \in \partial\Cinf$ 
which lies (approximately) on the boundary of the informed set (Sec.~\ref{subsec:grad})by applying Newton-Raphson method~\cite{RT06} (line~2).
Finally,~$x_0$ 
%use~$x_\text{boundary}$ 
is used (lines~3-5)
to initialize an MCMC algorithm which produces a series of samples within \Cinf (Sec.~\ref{subsec:mcmc}). Specifically, we chose to use HMC due to its favourable properties (see Sec.~\ref{sec:intro}). 
For a visualization of the approach, see Fig.~\ref{fig:alg}.

We apply MCMC by using the cost function to define a probability distribution \Pinf.
Specifically, points in \Cinf are distributed uniformly
while 
the probability of sampling any configuration $x \in \calX \setminus \Cinf$ is zero.
The MCMC algorithm will allow us to  efficiently produce samples (approximately) from \Pinf.

\begin{algorithm}[t]
	\begin{algorithmic}[1]
		\STATE $x_{\text{rand}} \leftarrow \texttt{sample\_uniform(\calX)}$
%		\STATE $x_{\text{boundary}} 
		\STATE $x_{0} 
		 \leftarrow
		 \texttt{newton\_raphson}(x_{\text{rand}}, c_{\text{best}})$
%		\STATE $x_{0} \leftarrow x_{\text{boundary}}$
		\FOR {$i = 1$ to $N$}
			\STATE $x_{i} \leftarrow \texttt{MCMC} (x_{i-1}, c_{\text{best}})$
			\STATE \textbf{output} $x_{i}$
		\ENDFOR

   	\end{algorithmic}
	\caption{MCMC-based Informed Sampling $(c_{\text{best}})$}
	\label{alg:mcmc_informed_sampling}	
\end{algorithm}

@@

@@

\begin{figure}[t!]
	\centering
	\begin{subfigure}[b]{0.5\textwidth}
		\includegraphics[width=\linewidth]{fig/sampling_efficiency/sample_efficiency_2d}
		\caption{2D}
		\label{fig:sampling_efficiency:2d}
	\end{subfigure}
	\begin{subfigure}[b]{0.5\textwidth}
		\includegraphics[width=\linewidth]{fig/sampling_efficiency/sample_efficiency_4d}
		\caption{4D}
		\label{fig:sampling_efficiency:4d}
	\end{subfigure}	
	\begin{subfigure}[b]{0.5\textwidth}
		\includegraphics[width=\linewidth]{fig/sampling_efficiency/sample_efficiency_6d}
		\caption{6D}
		\label{fig:sampling_efficiency:6d}
	\end{subfigure}
	\caption{\captionstyle Compare the sampling efficiency of different sampling methods.}
	\label{fig:sampling_efficiency}
\end{figure}

>>>>

\usepackage{xspace}
\usepackage{amsmath,amssymb,amsfonts,amsfonts}
\usepackage{algorithm}
\usepackage[noend]{algorithmic}
\usepackage{url}
\usepackage{bm}
\usepackage{graphicx}
\usepackage[font=footnotesize]{caption} % incompatible with aaai.sty?
\usepackage[font=footnotesize]{subcaption}
\usepackage{color}
\usepackage{cite}
\usepackage{hyperref}
\usepackage{eqparbox}
%\hypersetup{bookmarksopen,
%bookmarksnumbered,
%pdfpagemode=UseOutlines,
%colorlinks=true,
%linkcolor=blue,
%anchorcolor=blue,
%citecolor=blue,
%filecolor=blue,
%menucolor=blue,
%urlcolor=blue
%}

@@

\section{MCMC-based Informed Sampling}

In this section we describe our approach to efficiently produce a sequence of samples in the informed set~\Cinf given a specific cost~$c_{\text{best}}$.
Our algorithm (see Alg.~\ref{alg:mcmc_informed_sampling}) starts by randomly sampling a state~$x_\text{rand} \in \calX$ (line~2).
This sample is used to produces a 
sample~$x_0 \in \partial\Cinf$ 
%sample~$x_\text{boundary} \in \partial\Cinf$ 
which lies (approximately) on the boundary of the informed set (Sec.~\ref{subsec:grad}) by applying Newton-Raphson method~\cite{RT06} (line~3).
Finally,~$x_0$ 
%use~$x_\text{boundary}$ 
is used (lines~4-6)
to initialize an MCMC algorithm which produces a series of samples within \Cinf until some termination criterion is met.
This process is then repeated (Sec.~\ref{subsec:mcmc}). 
%Specifically, we chose to use HMC due to its favourable properties (see Sec.~\ref{sec:intro}). 
For a visualization of the approach, see Fig.~\ref{fig:alg}.

It is important to note that at each iteration, we will query our algorithm for one sample. However, this sample depends on the previous iterations of the algorithm which is why we describe it as a method to produce a sequence of samples.
Furthermore, the value~$c_{\text{best}}$ can decrease between consecutive iterations.
This will occur if the search algorithm that uses our sampler finds a path to the goal whose cost is lower than~$c_{\text{best}}$.

We apply MCMC by using the cost function to define a probability distribution \Pinf.
Specifically, points in \Cinf are distributed uniformly
while 
the probability of sampling any configuration $x \in \calX \setminus \Cinf$ is zero.
MCMC algorithms allow us to  efficiently produce samples (approximately) from \Pinf.

\begin{algorithm}[t]
	\begin{algorithmic}[1]
		\LOOP
			\STATE $x_{\text{rand}} \leftarrow \texttt{sample\_uniform(\calX)}$
			\STATE $x_{0} 
		 \leftarrow
		 \texttt{newton\_raphson}(x_{\text{rand}}, c_{\text{best}})$ ; \hspace{3mm} $i\leftarrow 0$
			\WHILE {\texttt{MCMC\_continue\_chain()}}
			\STATE $x_{i} \leftarrow \texttt{MCMC\_sample} (x_{i-1}, c_{\text{best}})$
			\STATE \textbf{output} $x_{i}$
			\ENDWHILE

		\ENDLOOP 
%		\STATE $x_{\text{rand}} \leftarrow \texttt{sample\_uniform(\calX)}$
%%		\STATE $x_{\text{boundary}} 
%		\STATE $x_{0} 
%		 \leftarrow
%		 \texttt{newton\_raphson}(x_{\text{rand}}, c_{\text{best}})$ ; \hspace{3mm} $i\leftarrow 0$
%%		\STATE $x_{0} \leftarrow x_{\text{boundary}}$
%		\WHILE {\texttt{MCMC\_continue\_chain()}}
%			\STATE $x_{i} \leftarrow \texttt{MCMC} (x_{i-1}, c_{\text{best}})$; \hspace{3mm} $i\leftarrow i+1$
%			\STATE \textbf{output} $x_{i}$
%		\ENDWHILE		
%		\FOR {$i = 1$ to $N$}
%			\STATE $x_{i} \leftarrow \texttt{MCMC} (x_{i-1}, c_{\text{best}})$
%			\STATE \textbf{output} $x_{i}$
%		\ENDFOR
%		 
   	\end{algorithmic}
	\caption{MCMC-based Informed Sampling $(c_{\text{best}})$}
	\label{alg:mcmc_informed_sampling}	
\end{algorithm}

@@

Our approach is general and can be applied to any MCMC algorithm (see Sec.~\ref{sec:related_work}).
We demonstrate how to instantiate it with two different algorithms: 
Hit and Run and Metropolis-Hastings.

For each algorithm, 
one has to describe how it is used to (i)~sample points 
(\texttt{MCMC\_sample, line 5})
and 
(ii)~decide if to continue sampling along the Markov Chain or to restart the entire process 
(\texttt{MCMC\_continue\_chain, line 5}).

\subsubsection{Hit and Run}

\subsubsection{Metropolis-Hasting}

\begin{algorithm}[t]
	\begin{algorithmic}[1]
		\STATE $ x'_{i+1} \leftarrow g( x'_{i+1} \mid x_{i} ) $
		\COMMENT {$g$~is a Gaussian}		
		\IF{ \textsc{Cost}($ x'_{i+1} $) $ < c_{\text{best}} $  }
		  \STATE $ x_{i+1} \leftarrow x'_{i+1} $ 
		\ELSE
		  \STATE $ x_{i+1} \leftarrow x_{i} $
		\ENDIF 
	\end{algorithmic}
	\caption{Metropolis-Hastings MCMC $(x_{i}, c_{\text{best}})$}
	\label{alg:mh_mcmc}	
\end{algorithm}

\begin{algorithm}[t]
	\begin{algorithmic}[1]
		\STATE $d \leftarrow$ \texttt{sample\_random\_direction}$()$
		\STATE $ L(\lambda) = \{  x \mid x = x_i + \lambda d_i \} $
		\STATE $ \lambda^{+} \leftarrow \sup L(\lambda) $; 
				\hspace{3mm} 
			   $ \lambda^{-} \leftarrow \inf L(\lambda) $

		\LOOP

		\STATE $ \lambda' \leftarrow$ \texttt{sample\_random}$ (\lambda^{-} , \lambda^{+})$
		\STATE $ x_{i+1} \leftarrow x_{i} + \lambda'_{i} d_i $

		\IF{ \textsc{Cost}($ x_{i+1} $) $ < c_{\text{best}} $ }
		  \RETURN $ x_{i+1}$
		\ENDIF

		\IF{ $ \lambda' > 0 $ }
			\STATE $ \lambda^{+} \leftarrow \lambda' $
		\ELSE
			\STATE $ \lambda^{-} \leftarrow \lambda'$
		\ENDIF

		\ENDLOOP
  	\end{algorithmic}
	\caption{Hit-And-Run MCMC $(x_{i}, c_{\text{best}})$}
	\label{alg:hit_and_run_mcmc}	
\end{algorithm}

@@

\begin{figure}[t!]
	\centering
	\begin{subfigure}[b]{0.5\textwidth}
		\includegraphics[width=\linewidth]{fig/sampling_efficiency/sample_efficiency_2d}
		\caption{2D}
		\label{fig:sampling_efficiency:2d}
	\end{subfigure}
	\begin{subfigure}[b]{0.5\textwidth}
		\includegraphics[width=\linewidth]{fig/sampling_efficiency/sample_efficiency_3d}
		\caption{3D}
		\label{fig:sampling_efficiency:4d}
	\end{subfigure}	
	\begin{subfigure}[b]{0.5\textwidth}
		\includegraphics[width=\linewidth]{fig/sampling_efficiency/sample_efficiency_6d}
		\caption{6D}
		\label{fig:sampling_efficiency:6d}
	\end{subfigure}
	\caption{\captionstyle Compare the sampling efficiency of different sampling methods.}
	\label{fig:sampling_efficiency}
\end{figure}
